---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```
# bayestools

## Overview

An R package for pre- and post-processing in phylogenetic and divergence-times bayesian inference. You can use it for processing output of phylogenetic bayesian programs (such as [Beast](https://www.beast2.org/) or [MrBayes](http://nbisweden.github.io/MrBayes/index.html)) or to prepare input files and decision making. It also supports sensitivity calculations using intersection (Ballen, in prep.) for comparing two given densities (e.g. prior vs. posterior).

## Installation

For now only the github development version is available. You can install it using `devtools`:

```{r, eval = FALSE}
install.packages("devtools")
devtools::install_github("gaballench/bayestools")
```

Once the package makes its way into CRAN the standard `install.packages` may be used for installing stable versions.

## Problems?

If you find a bug or unexpected behavior, please [file an issue](https://github.com/gaballench/bayestools/issues).

## Implementation and miscelaneous notes

### Why to use the median of quantiles as starting values?

on the initial values...
the following code will generate a lot of parameter estimates from several trials using different initial values. With a bit of large number theory, a decent estimate can be found.
Here, q and p are quantile and probabilities under a given distribution, a different anonymous function where sapply varies the values of the initial values allows to get the $par element of the optim call, and the the density plot shows that overall a median estimate approaches at the correct value. The values of q come from qDIST given the probabilities of interest (generally 0.05, 0.5, and 0.95). For instance: qbeta(p = c(0.05, 0.5, 0.95), shape1 = 10, shape2 = 1) for the example below:

```{r, echo = FALSE}
findParams <- function(q = c( -1.644854, 0, 1.644854), p = c(0.05,  0.50, 0.95), output = "complete", densit="pnorm", params = c("mean", "sd"), initVals = rep(1, times = l)) {
    l <- length(params)
    cl <- vector("list", 2  + length(params))
    cl[[1]] <- as.name(densit)
    cl[[2]] <- q
    names(cl) <- c(NA, "q", params)
    mode(cl) <- "call"
    quadraticFun <- function(x) {
        cl[3:(l+2)] <- x
        res <- eval(cl)
        sum((res - p)^2)
    }    
    res <- optim(initVals, quadraticFun)
    if (output == "parameters") {
        return(res$par)
    }
    return(res)
}
```

```{r, eval = TRUE, warning = FALSE}
# X = seq... is the set of values to try, from min(q) to max(q). 
parameters <- sapply(X = seq(0.7411344, 0.9948838, length.out = 1000),
                     FUN = function(x) {
                         findParams(q = c(0.7411344, 0.9330330, 0.9948838),
                                    p = c(0.05, 0.5, 0.95),
                                    densit = "pbeta",
                                    params = c("shape1", "shape2"),
                                    initVals = c(x, x))$par
                     },
                     simplify = TRUE)

plot(density(t(parameters)[, 1]))
plot(density(t(parameters)[, 2]))
```